Immune-Related Adverse Events (irAEs) associated with Immune Checkpoint Inhibitor (ICI) Therapy in Bladder Cancer: An Analysis of the FDA Adverse Event Reporting System (FAERS)

### 2025-03-14 code/eda-code/eda.qmd
Part 3 of MADA Project

Showing a bit of code to produce and save a summary table.
First, we summarize all our information using a table.

```{r}
library(dplyr)
library(tidyr)
library(kableExtra)

# Create a table for Age summary with Mean and SD
age_summary <- data.frame(
    Category = "Age",
    Value = paste(round(mean(mydata$patient_age, na.rm = TRUE), 2), "±", round(sd(mydata$patient_age, na.rm = TRUE), 2)),
    Count = NA,
    Percent = NA
)

# Sex Summary
sex_summary <- mydata %>%
    group_by(sex) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = sex) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Drug Summary
drug_summary <- mydata %>%
    group_by(suspect_product_active_ingredients) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = suspect_product_active_ingredients) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Outcome Summary
outcome_summary <- mydata %>%
    group_by(serious) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = serious) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Reporter Type Summary
reporter_summary <- mydata %>%
    group_by(reporter_type) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = reporter_type) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Country Summary
country_summary <- mydata %>%
    group_by(country_where_event_occurred) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = country_where_event_occurred) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Reaction Type Summary
# Define the organ system columns
organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")

summary_data <- mydata %>%
    select(all_of(organ_systems)) %>%
    pivot_longer(cols = everything(), values_to = "organ_system") %>%
    filter(!is.na(organ_system)) %>%
    group_by(organ_system) %>%
    summarise(
        Count = n(),
        Percent = round((n() / nrow(mydata)) * 100, 1)
    ) %>%
    rename(Category = organ_system) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Combine all summaries into one table
final_summary_table <- bind_rows(
    data.frame(Category = "Age", Value = age_summary$Value, stringsAsFactors = FALSE),
    data.frame(Category = "Sex", Value = "", stringsAsFactors = FALSE), # Subheading for Sex
    sex_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Sex
    data.frame(Category = "Immune Checkpoint Inhibitors", Value = "", stringsAsFactors = FALSE), # Subheading for Drugs
    drug_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Drugs
    data.frame(Category = "Outcome", Value = "", stringsAsFactors = FALSE), # Subheading for Outcome
    outcome_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Outcome
    data.frame(Category = "Reporter Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reporter Type
    reporter_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Reporter Type
    data.frame(Category = "Country", Value = "", stringsAsFactors = FALSE), # Subheading for Country
    country_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Country
    data.frame(Category = "Reaction Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reaction Type
    summary_data %>% select(Category, Value)
)

# Create a publication-quality table using kableExtra
final_summary_table %>%
    kable("html", caption = "Table 1: Descriptive Statistics", col.names = c("Category", "Value")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
    column_spec(1, bold = TRUE) %>%
    row_spec(which(final_summary_table$Category %in% c("Age", "Sex", "Immune Checkpoint Inhibitors", "Outcome", "Reporter Type", "Country", "Reaction Type")), bold = TRUE, background = "#f7f7f7") %>%
    footnote(general = "This table summarizes the patient data by various categories.")

@@

```{r}
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(final_summary_table, file = summarytable_file)
```

@@

```{r}
p2 <- summary_data %>%
    ggplot(aes(x = reorder(Category, Count), y = Count)) + # Reorder by Count ascending
    geom_col(fill = "#1f77b4") +
    labs(
        title = "Bar Plot of Reaction Types",
        x = "Organ System",
        y = "Count",
        caption = "Percentages are shown in parentheses"
    ) +
    theme_minimal() +
    geom_text(aes(label = Value), hjust = -0.1, size = 3) + # Adjust label position
    coord_flip() # Flip the axes

>>>>

First, we summarize all our information using a table.

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(huxtable)

# 1. Summarize Age (Mean and SD)
age_summary <- mydata %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(
    mean_age = mean(patient_age, na.rm = TRUE),
    sd_age = sd(patient_age, na.rm = TRUE)
  ) %>%
  mutate(`Age, mean years ± SD` = paste0(round(mean_age, 1), " ± ", round(sd_age, 1))) %>%
  select(suspect_product_active_ingredients, `Age, mean years ± SD`)

# 2. Summarize Sex Distribution
sex_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, sex) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = sex, values_from = count, values_fill = list(count = 0)) %>%
  rename(`Not Specified (Sex)` = `Not Specified`) %>%
  select(suspect_product_active_ingredients, Male, Female, `Not Specified (Sex)`)

# 3. Summarize Reaction Severity
reaction_severity_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, serious) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = serious, values_from = count, values_fill = list(count = 0)) %>%
  rename(Serious = Serious, `Non-Serious (Reaction)` = `Non-Serious`) %>%
  select(suspect_product_active_ingredients, Serious, `Non-Serious (Reaction)`)

# 4. Summarize Outcomes
outcomes_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, outcomes) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = outcomes, values_from = count, values_fill = list(count = 0)) %>%
  rename(`Non-Serious (Outcome)` = `Non-Serious`) %>%
  select(suspect_product_active_ingredients, Died, Hospitalized, Disabled, `Life Threatening`, `Non-Serious (Outcome)`, `Other Outcomes`)

# 5. Summarize Reporter Type
reporter_type_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, reporter_type) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = reporter_type, values_from = count, values_fill = list(count = 0)) %>%
  select(suspect_product_active_ingredients, `Healthcare Professional`, Consumer)

# 6. Summarize Country
country_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, country_where_event_occurred) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = country_where_event_occurred, values_from = count, values_fill = list(count = 0)) %>%
  rename(`Not Specified (Country)` = `Not Specified`) %>%
  select(suspect_product_active_ingredients, US, `Outside US`, `Not Specified (Country)`)

# Combine all summaries into one table
combined_summary <- age_summary %>%
  full_join(sex_summary, by = "suspect_product_active_ingredients") %>%
  full_join(reaction_severity_summary, by = "suspect_product_active_ingredients") %>%
  full_join(outcomes_summary, by = "suspect_product_active_ingredients") %>%
  full_join(reporter_type_summary, by = "suspect_product_active_ingredients") %>%
  full_join(country_summary, by = "suspect_product_active_ingredients")

# Convert all columns (except `suspect_product_active_ingredients`) to character
combined_summary <- combined_summary %>%
  mutate(across(-suspect_product_active_ingredients, as.character))

# Transpose the table for Word compatibility
transposed_table <- combined_summary %>%
  pivot_longer(cols = -suspect_product_active_ingredients, names_to = "Attribute", values_to = "Value") %>%
  pivot_wider(names_from = suspect_product_active_ingredients, values_from = Value)

# Rename columns for clarity
colnames(transposed_table)[1] <- "Attribute"

# Convert to huxtable object
huxtable_table <- as_hux(transposed_table)

# Format the table with huxtable (style, borders, etc.)
huxtable_table <- huxtable_table %>%
  set_all_borders(1) %>%
  set_font_size(12) %>%
  set_align("center") %>%
  set_width(1.5)

# Print the huxtable object, this will render correctly in Word
huxtable_table

@@

```{r}
# save to file
huxtable_table_path = here("results", "tables", "summarytable.rds")
saveRDS(huxtable_table, file = huxtable_table_path)
```

Table 2 - Summarizing organ system reaction by drug type
```{r}
# Load necessary libraries
library(huxtable)
library(dplyr)
library(tidyr)

# Define the organ system columns
organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")

# Transform the organ system columns into a single column and calculate counts and percentages
organ_system_summary <- mydata %>%
    select(suspect_product_active_ingredients, all_of(organ_systems)) %>%
    pivot_longer(cols = all_of(organ_systems), names_to = "Organ_System", values_to = "Reaction") %>%
    filter(!is.na(Reaction)) %>%
    group_by(suspect_product_active_ingredients, Reaction) %>%
    summarise(Count = n(), .groups = 'drop') %>%
    mutate(Percent = round((Count / nrow(mydata)) * 100, 1)) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Pivot the summary to have reactions as rows and drugs as columns, combining counts and percentages
organ_system_pivoted <- organ_system_summary %>%
    select(suspect_product_active_ingredients, Reaction, Value) %>%
    pivot_wider(names_from = suspect_product_active_ingredients, values_from = Value, values_fill = list(Value = "0"))

# Create a huxtable table
ht <- huxtable::as_hux(organ_system_pivoted)

# Add styling to the table
ht <- ht %>%
    set_bold(1, 1:ncol(ht)) %>%  # Bold the first row (header)
    set_caption("Immune Checkpoint Inhibitors by Organ System and Reaction") %>%
    set_width(1)                  # Set column width to make the table more compact

# Display the table
ht

#Save the table
ht_path = here("results", "tables", "summarytable1.rds")
saveRDS(ht, file = ht_path)

@@

```{r}
# Define the organ system columns
organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")

# Create a summary dataset for reactions across all organ systems
organ_system_summary <- mydata %>%
    select(all_of(organ_systems)) %>%
    pivot_longer(cols = everything(), names_to = "Organ_System", values_to = "Reaction") %>%
    filter(!is.na(Reaction)) %>%
    group_by(Reaction) %>%
    summarise(Count = n(), .groups = 'drop') %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# View the summary dataset
organ_system_summary

# Plot the reaction types with counts and percentages
p2 <- organ_system_summary %>%
    ggplot(aes(x = reorder(Reaction, Count), y = Count)) +  # Reorder by Count
    geom_col(fill = "#1f77b4") +
    labs(
        title = "Bar Plot of Reaction Types",
        x = "Reaction Type",
        y = "Count",
        caption = "Percentages are shown in parentheses"
    ) +
    theme_minimal() +
    geom_text(aes(label = Value), hjust = 1.1, size = 3) +  # Adjust label position to avoid overlap
    coord_flip() +  # Flip the axes to make it horizontal
    theme(
        plot.margin = margin(10, 50, 10, 10)  # Add margin to the right side to avoid label cut-off
    )

### 2025-03-14 code/processing-code/processingfile-v1.qmd
Part 3 of MADA Project

ici_data_subset <- ici_data1 %>%
  filter(str_detect(tolower(reactions), paste(tolower(reactions_of_interest), collapse = "|")))

@@

We have 368 reactions.
The next step is to characterize and categorize the irAEs.
The first step is to split the reaction column, and keep only the irAEs.

@@

@@

Let us classify the countries into US and Outside US
```{r}
ici_data_final <- ici_data_final %>%
  mutate(country_where_event_occurred = ifelse(country_where_event_occurred != "US", "Outside US", country_where_event_occurred))

@@

@@

# Define lists of reaction keywords for each organ system
dermatological_reactions <- c("rash", "pruritus", "dermatitis", "psoriasis")
gastrointestinal_reactions <- c("colitis", "hepatitis", "pancreatitis", "enterocolitis", "ileitis", "ileocolitis")
neurological_reactions <- c("myasthenia", "graves", "myelitis", "encephalitis", "neuritis", "peripheral neuropathy")
endocrinological_reactions <- c("hyperthyroidism", "hypothyroidism", "thyroiditis", "thyrotoxicosis", "adrenitis", "adrenal", "hypogonadism", "hyperparathyroidism", "hypercortisolism", "hypophysitis", "hypoparathyroidism", "hypocortisolism", "autoimmune diabetes")
cardiovascular_reactions <- c("myocarditis", "pericarditis")

# Assuming your data is stored in 'ici_data_final' and reactions are in reaction_1 to reaction_5
ici_data_final <- ici_data_final %>%
  mutate(
    organ_system_1 = case_when(
      !is.na(reaction_1) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_1) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_1) & grepl(paste(neurological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_1) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_1) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_1, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_1) ~ "Other",  # If not NA and no match, classify as Other
      TRUE ~ NA_character_  # If NA, keep as NA
    ),

    organ_system_2 = case_when(
      !is.na(reaction_2) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_2) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_2) & grepl(paste(neurological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_2) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_2) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_2, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_2) ~ "Other",
      TRUE ~ NA_character_
    ),

    organ_system_3 = case_when(
      !is.na(reaction_3) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_3) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_3) & grepl(paste(neurological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_3) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_3) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_3, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_3) ~ "Other",
      TRUE ~ NA_character_
    ),

    organ_system_4 = case_when(
      !is.na(reaction_4) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_4) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_4) & grepl(paste(neurological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_4) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_4) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_4, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_4) ~ "Other",
      TRUE ~ NA_character_
    ),

    organ_system_5 = case_when(
      !is.na(reaction_5) & grepl(paste(dermatological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Skin",
      !is.na(reaction_5) & grepl(paste(gastrointestinal_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Gastrointestinal",
      !is.na(reaction_5) & grepl(paste(neurological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Neurological",
      !is.na(reaction_5) & grepl(paste(endocrinological_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Endocrinological",
      !is.na(reaction_5) & grepl(paste(cardiovascular_reactions, collapse = "|"), reaction_5, ignore.case = TRUE) ~ "Cardiovascular",
      !is.na(reaction_5) ~ "Other",
      TRUE ~ NA_character_
    )
  )

@@

@@

>>>>

#Reactions with irAES
ici_data_subset <- ici_data1 %>%
  filter(str_detect(tolower(reactions), paste(tolower(reactions_of_interest), collapse = "|")))

#Reactions with irAEs
ici_data_non_irae <- ici_data1 %>%
  filter(!str_detect(tolower(reactions), paste(tolower(reactions_of_interest), collapse = "|")))

@@

We have 369 reactions.
The next step is to characterize and categorize the irAEs.
The first step is to split the reaction column, and keep only the irAEs.

@@

The next step is to remove any duplicates. We will remove reports that have identical age, weight, FDA received date, country of occurrence, suspected active ingredient, and reported reaction.

```{r}
# Find duplicates based on specified columns
duplicates <- ici_data_cleaned %>%
  group_by(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred) %>%
  filter(n() > 1) %>%
  ungroup()

# Display the duplicates
print(duplicates)
```

We see that there are 4 duplicates of 1 case.
We remove these.

```{r}
ici_data_cleaned = ici_data_cleaned %>%
  distinct(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred, .keep_all = TRUE)
```

@@

Let us classify the countries into US and Outside US
```{r}
ici_data_final <- ici_data_final %>%
  mutate(country_where_event_occurred = case_when(
    country_where_event_occurred == "US" ~ "US",
    country_where_event_occurred != "US" & country_where_event_occurred != "Not Specified" ~ "Outside US",
    TRUE ~ country_where_event_occurred
  ))

@@

The next step is to classify the irAEs by the System Organ Class (SOC) according to MedDRA.

@@

# Define lists of reaction keywords for each organ system
Skin_and_subcutaneous_tissue_disorders <- c("rash", "pruritus", "dermatitis", "psoriasis")
Gastrointestinal_disorders <- c("colitis", "hepatitis", "pancreatitis", "enterocolitis", "ileitis", "ileocolitis")
Nervous_system_disorders <- c("myasthenia", "graves", "myelitis", "encephalitis", "neuritis", "peripheral neuropathy")
Endocrine_disorders <- c("hyperthyroidism", "hypothyroidism", "thyroiditis", "thyrotoxicosis", "adrenitis", "adrenal", "hypogonadism", "hyperparathyroidism", "hypercortisolism", "hypophysitis", "hypoparathyroidism", "hypocortisolism", "autoimmune diabetes")
Cardiac_disorders <- c("myocarditis", "pericarditis", "myocarditis")
Respiratory_thoracic_and_mediastinal_disorders <- c("pneumonitis")
Renal_and_urinary_disorders <- c("pyelonephritis", "nephritis")

# Function to classify reactions into organ systems
classify_reaction <- function(reaction) {
  case_when(
    !is.na(reaction) & grepl(paste(Skin_and_subcutaneous_tissue_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Skin",
    !is.na(reaction) & grepl(paste(Gastrointestinal_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Gastrointestinal",
    !is.na(reaction) & grepl(paste(Nervous_system_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Neurological",
    !is.na(reaction) & grepl(paste(Endocrine_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Endocrinological",
    !is.na(reaction) & grepl(paste(Cardiac_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Cardiovascular",
    !is.na(reaction) & grepl(paste(Respiratory_thoracic_and_mediastinal_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Respiratory",
    !is.na(reaction) & grepl(paste(Renal_and_urinary_disorders, collapse = "|"), reaction, ignore.case = TRUE) ~ "Renal",
    !is.na(reaction) ~ "Other",  # If not NA and no match, classify as Other
    TRUE ~ NA_character_  # If NA, keep as NA
  )
}

# Apply the function to each reaction column
ici_data_final <- ici_data_final %>%
  mutate(
    organ_system_1 = classify_reaction(reaction_1),
    organ_system_2 = classify_reaction(reaction_2),
    organ_system_3 = classify_reaction(reaction_3),
    organ_system_4 = classify_reaction(reaction_4),
    organ_system_5 = classify_reaction(reaction_5)
  )

@@

```
Now, let us split the outcomes column.

ici_data_final <- ici_data_final %>%
  separate(outcomes, into = paste0("outcome_", 1:2), sep = ";", remove = FALSE, extra = "drop")

# If outcome is died, removing other outcomes for that report
ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Died"] <- NA
ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Died"] <- NA

#If outcome is hospitalized, removing other outcomes for that report
ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Hospitalized"] <- NA
ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Hospitalized"] <- NA

#If outcome is life threatening, removing other outcomes for that report
ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Life Threatening"] <- NA
ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Life Threatening"] <- NA

#If outcome is disabled, removing other outcomes for that report
ici_data_final$outcome_1[!is.na(ici_data_final$outcome_1) & ici_data_final$outcome_2 == "Disabled"] <- NA
ici_data_final$outcome_2[!is.na(ici_data_final$outcome_2) & ici_data_final$outcome_1 == "Disabled"] <- NA

#Merge the columns
ici_data_final$outcomes <- coalesce(ici_data_final$outcome_1, ici_data_final$outcome_2)  # Merge outcome_1 and outcome_2
ici_data_final$outcome_1 <- NULL  # Remove outcome_1 column
ici_data_final$outcome_2 <- NULL  # Remove outcome_2 column

# View the result
View(ici_data_final)

@@

Since we will need the number of reactions other than irAEs for the further analysis, we summarize that information here.

```{r}
glimpse(ici_data_non_irae)
```

First, let us split the reactions column
```{r}
library(tidyr)

ici_data_non_irae <- ici_data_non_irae %>%
  separate(reactions, into = paste0("reaction_", 1:15), sep = ";", remove = FALSE, extra = "drop")

# View the result
View(ici_data_non_irae)

reaction_columns <- paste0("reaction_", 1:15)
```

We remove any duplicates.
```{r}
duplicates1 <- ici_data_non_irae %>%
  group_by(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred) %>%
  filter(n() > 1) %>%
  ungroup()

# Display the duplicates
print(duplicates1)
```

We see multiple duplicates. Let us remove them.
```{r}
ici_data_clean_non_irae = ici_data_non_irae %>%
  distinct(suspect_product_active_ingredients, sex, patient_age, patient_weight, reactions, initial_fda_received_date, country_where_event_occurred, .keep_all = TRUE)

View(ici_data_clean_non_irae)
```

We will need this data in the disproportionality analysis.

We save this data as well.

```{r}
processeddata1 = ici_data_clean_non_irae
save_data_location <- here::here("data","processed-data","processeddata1.rds")
saveRDS(processeddata1, file = save_data_location)
```

### 2025-03-14 products/manuscript/Manuscript.qmd
Fixed the rendering error
Part 3 of MADA Project

*Write a summary of your project.* *This part has not been completed. The source code has been left as is*

@@

*In order to obtain the FAERS data, the ICIs were used to search by product. Pembrolizumab, atezolizumab, avelumab, and nivolumab were the keywords used. Previously, durvalumab, and ipilimumab were also approved for bladder cancer treatment. However, this indication has since then been removed by the manufacturers. As such, these were not considered. Following this, the search was limited to cases of bladder cancer. A total of 1,886 reactions were obtained.*

@@

*Reactions with concomitant suspect products, combination products, and those which were not irAEs were removed. This resulted in a total of 369 cases. The details are available in the processing code file.*

@@

*The second part of the analysis will involve a disproportionality analysis. This analysis uses the following formula to calculate the reported odds ratio (ROR).*

*a = irAES reported with ICIs* *b = irAEs reported with all other drugs* *c = All other AEs reported with ICIs* *d = All other AEs reported with other drugs*

*ROR = a/c / b/d*

*Other details in the analysis will be finalized later.*

@@

*This part has not been completed. The source code has been left as is*

@@

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Table 1: Descriptive Statistics"
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

@@

@@

*This part has not yet been edited.*

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

>>>>

*Background: Bladder cancer is the ninth most common type of cancer worldwide. Treatment options include surgery, chemotherapy and immunotherapy. Immune checkpoint inhibitors (ICIs) are a class of immunotherapeutic agents recommended for the treatment of bladder cancer. ICIs have been associated with certain immune-related adverse events (irAEs)*

*Methods: Reports in the FAERS dashbaord through December 2024 were analyzed for irAEs to four ICIs recommended in bladder cancer - atezolizumab, avelumab, nivolumab, and pembrolizumab. Standardized queries were used to identify cases of rash, pruritus, hypothyroidism, hyperthyroidism, colitis, hepatitis, and nephritis. Descriptive and disproportionality analyses were performed to detect safety signals.*

*Results: To be added in Part 4*

@@

*In order to obtain the FAERS data, the ICIs were used to search by product. Pembrolizumab, atezolizumab, avelumab, and nivolumab were the keywords used. Previously, durvalumab, and ipilimumab were also approved for bladder cancer treatment. However, this indication has since then been removed by the manufacturers. As such, these were not considered. Following this, the search was limited to cases of bladder cancer. A total of 2948 reactions were obtained.*

@@

*Reactions with concomitant suspect products, combination products, and those which were not irAEs were removed. This resulted in a total of 365 cases. The details are available in the processing code file.*

*The processing was done according to @fig-processing1.*

```{r}
#| label: fig-processing1
#| fig-cap: "Study Flowchart"
#| echo: FALSE
knitr::include_graphics(here("assets","Figure1.png"))
```

@@

*The second part of the analysis will involve a disproportionality analysis. This analysis employs Proportional Reporting Ratio (PRR), Reporting Odds Ratio (ROR), and Information Component (IC). Further details will be added in Part 4..*

@@

@@

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Table 1: Descriptive Statistics"
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)

```

```{r}

```

@@

*@tbl-summarytable1 shows the number of reactions by organ system and the ICI.*

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "Table 2: Reactions by organ system and drug"
#| echo: FALSE
resulttable1=readRDS("../../results/tables/summarytable1.rds")
knitr::kable(resulttable1)
```

@@

*The following tables show the results from the disproportionality analysis.*

```{r}
#| label: tbl-atezolizumab
#| tbl-cap: "Table 1: Atezolizumab Reactions by Organ System"
#| echo: FALSE
table_atezolizumab <- readRDS("../../results/tables/table_atezolizumab.rds")
knitr::kable(table_atezolizumab)
```

```{r}
#| label: tbl-avelumab
#| tbl-cap: "Table 2: Avelumab Reactions by Organ System"
#| echo: FALSE
table_avelumab <- readRDS("../../results/tables/table_avelumab.rds")
knitr::kable(table_avelumab)
```

```{r}
#| label: tbl-nivolumab
#| tbl-cap: "Table 3: Nivolumab Reactions by Organ System"
#| echo: FALSE
table_nivolumab <- readRDS("../../results/tables/table_nivolumab.rds")
knitr::kable(table_nivolumab)
```

```{r}
#| label: tbl-pembrolizumab
#| tbl-cap: "Table 4: Pembrolizumab Reactions by Organ System"
#| echo: FALSE
table_pembrolizumab <- readRDS("../../results/tables/table_pembrolizumab.rds")
knitr::kable(table_pembrolizumab)

```

*The PRR and ROR when all irAEs are considered is greater than 1 for pembrolizumab (PRR: 1.8 (1.52, 2.14), ROR: 2.07 (1.69, 2.55)). Similarly, the IC is greater than 0 (0.36 (0.29, 0.44)). Thus, there is a safety signal associated with pembrolizumab for all irAEs.*

*When considering individual irAEs, pembrolizumab also has a safety signal for rash (PRR: 2.11 (1.17, 3.8), ROR: 2.14 (1.18, 3.9), IC: 0.44 (0.17, 0.71).*

*Atezolizumab has a safety signal for hyperthyroidism (PRR: 3.51 (1.07, 11.45), ROR: 3.53 (1.07, 11.61), IC:	1.1 (0.33, 1.87). There also may be a signal for nephritis (PRR: 3.65 (0.98, 13.56), ROR: 3.68 (0.98, 13.73), IC:	1.12 (0.29, 1.96), though the 95% CI for PRR and ROR is not greater than 1 (though very close to 1).*

*Similarly, avelumab may has a safety signal for hypothyroidism (PRR: 3.05 (0.98, 9.51), ROR:	3.08 (0.97, 9.75), IC:	1.32 (0.12, 2.52)).*

*No significant safety signals were found for nivolumab.*
